---
title: 
author: 
date: 
output: html_document
---
Gathering Data:
```{r}
heartmortCOMPLETE = read.csv(file = "https://raw.githubusercontent.com/PDBoegel/HeartMortality/master/HeartMort.csv", header = TRUE)
heartmort = heartmortCOMPLETE[-c(1:4,7,24:26)] #Getting rid of variables for metadata and other dependent variable
options(max.print = 4000)
```

# Basic data manipulation to better fit normality assumptions for linear models

```{r}
heartmort$PCTPOPUNDER65 = heartmortCOMPLETE$PCTPOP30TO65 + heartmortCOMPLETE$PCTPOPUNDER30 # To account for elderly population
heartmort <- subset(heartmort, AVGINCOMEPERPERSON > 0 & PCTASIANAM != 0 & PCTAFRAM != 0 & PCTNATIVEAM != 0) # 26 counties had 0 avgincome/PCTAFRAM/PCTNATIVEAM/PCTASAINAM, removal should not affect study and makes histogram/log transform variables easier to make/use later
summaryOfVars <- summary(heartmort)
summaryOfVars
```
```{r, echo=FALSE}
heartmort <- heartmort[,c(1:6,14:16,19,7:13,18,20)] #Organization to make category variables easier later on. Does not need to go into document

heartmort[,c(7:19)] <- heartmort[,c(7:19)]/100 # Making percents into numbers [0,1]
```

```{r, results='hide', echo=FALSE}
# Put These into the appendix
# par(mfrow = c(2,3))
# lapply(1:ncol(heartmort), function(i) hist(heartmort[,i], xlab = names(heartmort)[i], main = paste("Histogram of", names(heartmort)[i])))
```

From looking at the summary and histograms of the Variables, those that need to be transformed are:

Log transform:
POPSIZE [log]
HDMORTPERPERSON [log]
POPDENSITY [log]
HOUSINGUNITS [log]
HOUSINGUNITDENSITY [log]
AVGINCOMEPERPERSON [log]
TOTUNEMPLOYED [log]

categorized:
All Percentage variables besides diabetes and obesity

Notes: Some data in AVGINCOMEPERPERSON is zero which is strange might want to consider removing these points, otherwise we add one so their log values are inf    
```{r, results='hide'}
# We log totals imeediately as popsize varies widely from county to county
totCols <- c("LNTOTLOWEDU","LNTOTMEDEDU","LNTOTHIGHEDU","LNTOTOBESE","LNTOTDIABETES","LNTOTMALE","LNTOTFEMALE","LNTOTWHITE","LNTOTAFRAM","LNTOTASIANAM","LNTOTNATIVEAM","LNTOTUNEMPLOYED","LNTOTNONELDERLY")
heartmort[totCols] <- log(heartmort$POPSIZE * heartmort[7:19])

# Making categories for PCT variables
catCols <- c("CATHDMORTPERPERSON","CATLOWEDU","CATMEDEDU","CATHIGHEDU","CATOBESE","CATDIABETES","CATMALE","CATFEMALE","CATWHITE","CATAFRAM","CATASIANAM","CATNATIVEAM","CATUNEMPLOYED","CATNONELDERLY")
heartmort[catCols] <- 0
df <- lapply(c(2,7:19), function(i) cut(heartmort[,i], c(-.01, max(0, as.numeric(quantile(heartmort[,i], probs = 0.25)) - (1.5 * (as.numeric(quantile(heartmort[,i], probs = 0.75)) - as.numeric(quantile(heartmort[,i],probs = 0.25))))), min(1, as.numeric(quantile(heartmort[,i], probs = 0.75)) + (1.5 * as.numeric((quantile(heartmort[,i], probs = 0.75)) - as.numeric(quantile(heartmort[,i], probs = 0.25))))),1.01), right = FALSE, labels = c(0:2)))

heartmort[33:46] <- df
colnames(heartmort)[33:45] <- catCols
# Log transformations from what was specified above
logCols <- c("LNPOPSIZE","LNHDMORTPERPERSON","LNPOPDENSITY","LNHOUSINGUNITS","LNHOUSINGUNITSDENSITY","LNAVGINCOMEPERPERSON")
heartmort[logCols] <- log(heartmort[1:6])
par(mfrow = c(2,3))
```

```{r, results='hide', echo=FALSE}
# Put these into the appendix
# par(mfrow = c(2,3))
# lapply(21:ncol(heartmort), function(i) hist(heartmort[,i], xlab = names(heartmort)[i], main = paste("Histogram of", names(heartmort)[i]))) # This goes into the appendix
```


**Table of Correlations:**
```{r}
#Table of correlations : EDUCATION LEVEL AND OBESITY/DIABETES RATES
cor(heartmort[20:24])

#Table of correlations : Demographic
cor(heartmort[25:30])

#Table of correlations : Population, Housing, Income
# cor(heartmort[c(46,48:51)])
```

**linear model analysis**

```{r}
attach(heartmort)
totlm = lm(LNHDMORTPERPERSON ~ LNTOTLOWEDU + LNTOTMEDEDU + LNTOTHIGHEDU + LNTOTAFRAM + LNTOTASIANAM + LNTOTDIABETES + LNTOTFEMALE + LNTOTMALE + LNTOTNATIVEAM + LNTOTNONELDERLY + LNTOTOBESE + LNTOTUNEMPLOYED + LNTOTWHITE + LNPOPSIZE + LNHOUSINGUNITS, data = heartmort)
summary(totlm)
```

```{r}
catlm = lm(LNHDMORTPERPERSON ~ LNPOPDENSITY + LNHOUSINGUNITSDENSITY + factor(CATAFRAM) + factor(CATASIANAM) + factor(CATDIABETES) + factor(CATFEMALE) + factor(CATHIGHEDU) + factor(CATLOWEDU) + factor(CATMALE) + factor(CATMEDEDU) + factor(CATNATIVEAM) + factor(CATNONELDERLY) + factor(CATOBESE) + factor(CATUNEMPLOYED) + factor(CATWHITE), data = heartmort)
summary(catlm)
```

**Sensitivity Analysis** 

```{r, eval = FALSE}
#CHANGE tlm BASE UPON WHICH MODEL CHOSEN FROM ABOVE
resid = rstandard(tlm)
leverage = hatvalues(tlm) 

hist(resid, main="Histogram of Standardized Residuals", xlab = "Standardized Residual")


k = dim(heartmort)[2]
n = dim(heartmort)[1]
averagelev = (k+1)/n
3*averagelev
hist(leverage, main="Histogram of Leverages", xlab = "Leverage", breaks = 15)
```


```{r, eval = FALSE}
which(leverage > 3*averagelev)
length(which(leverage > 3*averagelev))

which(leverage > 6*averagelev)
length(which(leverage > 6*averagelev))

```

```{r, eval = FALSE}
levs <-as.numeric(which(leverage > 3*averagelev))
levs
lev6heartmort = heartmort[-levs,]
lmlev6 = lm(HDMORTPERPERSON ~ ., data = lev6heartmort)
summary(lmlev6)
```

```{r, eval = FALSE}
residlev6 = rstandard(lmlev6)
leveragelev6 = hatvalues(lmlev6) 

hist(residlev6, main="Histogram of Standardized Residuals", xlab = "Standardized Residual")


k = dim(lev6heartmort)[2]
n = dim(lev6heartmort)[1]
averagelev = (k+1)/n
3*averagelev
hist(leveragelev6, main="Histogram of Leverages", xlab = "Leverage", breaks = 15)
```
*End sensitivity analysis*


### Notes ############################

Get rid of percents - multiply demographic percents by population in each county

Eliminate high leverage points before transforming variables


